A Continual Planner that Switches Between Serial and Decision-Theoretic Planning

### 2010-11-17 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
New notations.
 - Charles

We describe, \pcogx, the planning component in a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire a good
understanding of its environment.
%%

@@

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying world
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.

A problem state $\state$ is a set of propositions $\state \subseteq
\props$. Every stochastic action $\stochAction \in \stochActions$ has
a precondition $\poss(\stochAction)$, also a set of propositions. An
action is applicable at a state $\state$ when $\poss(\stochAction)
\subseteq \state$. We denote $\stochActions(\state)$ the set of
actions applicable at state $\state$.  When $\stochAction \in
\stochActions(\state)$ is applied, nature chooses one element amongst
a small set of deterministic outcomes $\detActions(\stochAction)
\equiv \{\detAction_1, \ldots, \detAction_k \}$. We denote
$\mu_{\stochAction}(\detAction_i)$ the probability that nature takes
outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. That outcome so chosen has an
effect on the underlying state given in terms of two lists of
propositions. The add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
  $\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
  ..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic senses $\stochSenses$, deterministic sensing
outcomes $\detSenses$, and perceptual propositions $\percepts$, called
percepts. Here, an observation $\observ$ is a set of percepts $\observ
\subseteq \percepts$, and we denote \observations\ the set of
reachable observations. The underlying state of the world cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in \observations$ that informs what
should be believed about the world. In the POMDP model, if $\action$
is applied effecting a transition to a successor state $\state'$, then
an observation occurs according to the active senses
$\stochSenses(\stochAction, \state') \subseteq \stochSenses$. A
precept is active, written $\stochSense \in \stochSenses(\stochAction,
\state')$, if the senses action-precondition,
$\poss_\stochActions(\stochSense)$, is equal to $\stochAction$, and
the state-precondition $\poss_\states(\stochSense) \subseteq
\propositions$ is satisfied by the state $\state'$; In the usual sense
that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a subjective
Bayesian belief state. Written $\bstate_0$, this is a probability
distribution over states. We use the notation $\bstate_0(\state)$ to
denote the probability that the starting state is equal to
$\state$. We suppose this is given in a factored tree format.

%% Let $\detSense_i \subseteq \percepts$ for each $i$, so that if
%% we apply $\stochAction$ and transition to $\state'$, the observation
%% received will be a union of choices from the active percepts, which
%% will occur with the probability according to $\psi$ that nature can
%% make those choices.

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
we do incorporate a PPDDL-like default semantics, supposing that when
an action \stochAction\ is executed at some \bstate, that it has no
effect on states $\bstate(\state) > 0$ where $\poss(\stochAction)
\not\subseteq \state$.

@@

Whereas until now we have considered a POMDP model that is factored in
terms of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations.  Where $\state, \state' \in
\states$, $a \in \actions$, from $\mu$ we have a state transition
function $\transProb(\state, \action, \state')$ giving the probability
of a transition from state $s$ to $s'$ if $a$ is applied. For any
$\state$ and $\action$ we have $\sum_{\state' \in \states}
\transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

@@

\begin{equation}\label{eq:evaluation}
V_\node(\state) = \sum_{\action \in \actions}
\selection_\node(\action) \reward(\state, \action) + \beta \sum_{\action, \observation,
\state', \node'} \transition_\node(\action, \observation, \node')
\transProb(\state, \action, \state') \obsDist_\observ(\state',
\action) V_{\node'}(\state')
\end{equation}

Using the notation \bstate\ for a belief-state, and writing
$\bstate(\state)$ for the probability the robot believes it is in
state \state, we have that the value of that belief $\bstate$ according to a
controller is:

>>>>

We describe, \pcogx, the planning component in a robotic system that
continuously deliberates in a stochastic dynamic environment in order
to achieve objectives set by the user, and acquire knowledge about its
surroundings.
%%

@@

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. The underlying process
dynamics are modelled in terms of a finite set of stochastic actions
$\stochActions$, deterministic outcomes $\detActions$, and
state-characterising propositions $\propositions$.

We model a process state $\state$ as the set of propositions that are
true of the state. Notationally, we have $\state \subseteq
\props$. State change is induced by application of actions. A
stochastic action $\stochAction \in \stochActions$ is applicable if
its precondition $\poss(\stochAction)$, a set of propositions, are
satisfied in the current state -- I.e., $\poss(\stochAction) \subseteq
\state$. We write $\stochActions(\state)$ for set of actions
applicable at state $\state$.  If action $\stochAction \in
\stochActions(\state)$ is applied at \state, nature chooses one
element amongst a small set of deterministic outcomes
$\detActions(\stochAction) \equiv \{\detAction_1, \ldots, \detAction_k
\}$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature takes outcome $\detAct_i$, and for all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$. The chosen outcome has an
effect on the state given in terms of two lists of propositions. The
add-list $\add(\detAction)$ and delete-list
$\delete(\detAction)$.\footnote{If a proposition is in the add-list of
$\detAction$, then it cannot be in the delete-list and vice versa.}
If outcome $\detAction$ with $\add(\detAction) := [\prop_1,
..,\prop_n]$ and $\delete(\detAction) := [\prop_1, ..,\prop_m]$ is
chosen by nature when \stochAction\ is applied at state $\state$, then
the resultant state is $ ( \state \cup \add(\detAction) ) \backslash
\delete(\detAction)$ -- i.e., propositions from $\add(\detAction)$ are
added to $\state$, and those from $\delete(\detAction)$ are removed
from $\state$.

We are concerned with problems that feature partial
observability. These have a perceptual model given in terms of a
finite set of stochastic {\em senses} $\stochSenses$, deterministic
sensing outcomes $\detSenses$, and perceptual propositions
$\percepts$, called {\em percepts}. Here, an observation $\observ$ is
a set of percepts $\observ \subseteq \percepts$, and we denote
\observations\ the set of reachable observations. The underlying state
of the process cannot be observed directly, rather, senses
$\stochSense \in \stochSenses$ effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. In detail, if $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A precept is active, written $\stochSense \in
\stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$; In the
usual sense that $\poss_\states(\stochSense) \subseteq \state'$.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from chosen elements of active senses.

A POMDP has a starting configuration that corresponds to a Bayesian
belief-state. Intuitively, this is the robot's subjective belief about
its environment. Technically, a belief-state $\bstate$ is a
probability distribution over process states. We write
$\bstate(\state)$ to denote the probability that the process is in
$\state$ according to $\bstate$, and $\bstate_0$ when discussing the
starting configuration. We suppose the latter is given in a factored
tree format.

Finally, we make a number of fairly standard assumptions. First, that
action execution and sensing occurs instantaneously, and that only one
action can be applied at a plan-step. Second, it can arise that in
some propositional states an action is applicable, and in others it is
not. Moreover, a belief-state can assign non-zero probability to
states where that applicability holds, and states where it does
not. We differ
from~\citeauthor{younes:littman:04}~\citeyear{younes:littman:04},
because we forbid the execution of actions that are not applicable in
any state of the current belief.  Also,
unlike~\citeauthor{hoffmann:brafman:2006}~\citeyear{hoffmann:brafman:2006},
we do incorporate a PPDDL-like default semantics, supposing that when
an action \stochAction\ is executed at $\state_i$ where
$\bstate(\state_i) > 0$, that it has no effect on states $\state_i$ if
$\poss(\stochAction) \not\subseteq \state_i$ and $\bstate(\state_i) >
0$.

@@

Whereas until now we have considered a POMDP model that is factored in
terms of propositions and percepts, in order to discuss utilities and
policies it is convenient to consider the underlying decision process
in a flat format. This underlying decision process is given by the
tuple $\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
Function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function.  Consequently there is a positive constant $c$
so that for all $\state \in \states$ and $\action \in \actions$,
$|\reward(\state, \action)| < c$.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

@@

\begin{equation}\label{eq:evaluation}
\begin{array}{lcl}
V_\node(\state) & = & \sum_{\action \in \actions}
\selection_\node(\action) \reward(\state, \action) \;\; + \vspace{1ex} \\

&& \hspace{-10ex} \beta \sum_{\action, \observation,
\state', \node'} \transition_\node(\action, \observation, \node')
\transProb(\state, \action, \state') \obsDist_\observ(\state',
\action) V_{\node'}(\state')
\end{array}
\end{equation}

\noindent The value of $\bstate$ according to a controller is
therefore:
